### 2017-11-30 MSDS_CaseStudy2_Master.Rmd
Jose and Lisa R files combined - Draft 1
We can begin working off this _Master file and merging into it.

---
title: "MSDS Case Study 2 - Procrastination and HDI Data"
author: "Jose Torres and Lisa Street"
date: "November 29, 2017"  
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing Necessary Libraries

Install necessary packages if not installed:

```{r custom packages}
if (!is.element("dplyr", installed.packages()[,1]))
  install.packages("dplyr",repos="http://cran.rstudio.com")
if (!is.element("ggplot2", installed.packages()[,1]))
  install.packages("ggplot2",repos="http://cran.rstudio.com")
if (!is.element("stringr", installed.packages()[,1]))
  install.packages("stringr",repos="http://cran.rstudio.com")
if (!is.element("rvest", installed.packages()[,1]))
  install.packages("rvest",repos="http://cran.rstudio.com")
if (!is.element("xml2", installed.packages()[,1]))
  install.packages("xml2",repos="http://cran.rstudio.com")
 if (!is.element("tidyr", installed.packages()[,1]))
  install.packages("tidyr",repos="http://cran.rstudio.com")
if (!is.element("curl", installed.packages()[,1]))
  install.packages("curl",repos="http://cran.rstudio.com")
if (!is.element("stargazer", installed.packages()[,1]))
  install.packages("stargazer",repos="http://cran.rstudio.com")
if (!is.element("stringi", installed.packages()[,1]))
  install.packages("stringi",repos="http://cran.rstudio.com")
```

```{r loadlibs, include=FALSE}

#Load libraries
library("dplyr")
library("rvest")
library("tidyr")
library("ggplot2")
library("stargazer")
library("ggpubr")
library("reshape2")
library("stringr")

```

## Load custom functions.

```{r custom functions}

#Custom functions for Case Study 2
#Last modified 11/26/17

#Wrapper function for gsub that maintains object classes
gsubkeep = function(pattern='',replacement=NA,x, ignore.case=FALSE, perl=FALSE,
										fixed=FALSE, useBytes=FALSE) {
	if (is.null(dim(x)) && !is.null(class(x))) {
		colObj = x
		colClass = class(colObj)
		colObj = gsub(pattern=pattern,replacement=replacement,x=colObj,
									ignore.case=ignore.case,perl=perl,fixed=fixed,
									useBytes=useBytes)
		if (colClass == 'factor') {
			colObj = as.factor(colObj)
		} else { 
			colObj = as(colObj,colClass) 
		}
		x = colObj
	} else {
		for (i in 1:length(x)) {
			colObj = x[,i]
			colClass = class(colObj)
			colObj = gsub(pattern=pattern,replacement=replacement,x=colObj,
										ignore.case=ignore.case,perl=perl,fixed=fixed,
										useBytes=useBytes)
			if (colClass == 'factor') {
				colObj = as.factor(colObj)
			} else { 
				colObj = as(colObj,colClass) 
			}
			x[,i] = colObj
		}
	}
	return(x)
}

#Function that row-stacks the identified columns of a list of data frames
#Parameters: {x: list of data frames, elements: the particular data frames within x to stack, 
#             cols: the particular columns to stack, colname: names of the columns selected with cols parameter }
list_rowbind = function(x,elements=1:length(x),cols = columns(1:length(x[[1]])),
												colname = colnames(x[[1]])[cols]) {
	result = data.frame()
	for (i in elements) {
		append_df = x[[i]][,cols]
		colnames(append_df) = colname
		result = bind_rows(result,append_df)
	}
	return(result)
}

#Function that creates a list of indices and corresponding strings which match keywords
#Paratermers: {x: a vector (to be coerced to character) to search, findtext: a vector of keywords to look for}
match_list = function(x,findtext='') {
	findtext = as.character(findtext)
	x = tolower(as.character(x))
	result = list(length(findtext))
	for (i in 1:length(findtext)) {
		result[i] = list(cbind('Idx'=grep(findtext[i],x),'Match'=x[grep(findtext[i],x)]))
	}
	return(result)
}

```

## Question 2: Read in and clean raw data

```{r Question2}

# 2: Read in and clean raw data

# 2A: Read in CSV
dataset_raw <- read.csv("C:\\SMU_INTRO_TO_DS\\CASESTUDY2\\Data\\Procrastination.csv")

dim(dataset_raw)
# There are 4264 rows and 61 columns in the procrastination dataset.

# 2B: Partition Procrastination into 2 tables (Attributes and Responses/Answers) and update column names
ID = 1:dim(dataset_raw)[1]
resp_attr = cbind(ID,dataset_raw[,1:14])
resp_answers = cbind(ID,dataset_raw[,15:length(dataset_raw)])

resp_attr_names = c('ID','Age','Gender','Kids','Edu','Work_Status',
										'Annual_Inc','Current_Occ','Employ_Yrs','Employ_Mths',
										'Comm_Size','Country_Res','Married','Num_sons',
										'Num_daught')

resp_answer_names = c('ID',paste('DP_',1:5,sep=''),paste('AIP_',1:15,sep=''),
											paste('GP_',1:20,sep=''),paste('SWLS_',1:5,sep=''),'Self_Score','Others_Score')

colnames(resp_attr) = resp_attr_names
colnames(resp_answers) = resp_answer_names

# 2C: Cleanup the data

# 2Ci through 2civ:

# Begin by converting all blank or empty strings to NA in the entire data set
resp_attr = gsubkeep(pattern='^$|^ $',replacement=NA,x=resp_attr)

# Now go through individual attributes list and replace 'bad' values with NA
cleanup_list = data.frame(rbind(c(' Kids$','',4), #remove 'Kids'
																c('0', NA, 6), #Work status
																c('0|please specify',NA,8), #Occupation 
																c('999',NA,9), #Employment Years
																c('[08]',NA,11), #Community Size
																c('0',NA,12), #Country of Residence
																c('0',NA,13), #Married
																c('Male','1',14), #Number of sons
																c('Female','2',14))) #Number of sons

# Employ_Yrs_num - replace small numbers with 0.  Round to nearest integer.
resp_attr[which(resp_attr[,9] < .5),9] = 0
resp_attr[,9] = as.integer(round(resp_attr[,9],digits=0))

# Num_sons_fac - Convert to numeric
resp_attr[,14] = as.integer(resp_attr[,14])

# Num_daughters_num - no change

# 2d:
# Set columns to the proper data types
for (i in 1:dim(cleanup_list)[1]) {
	pattern = as.character(cleanup_list[i,1])
	replacement = as.character(cleanup_list[i,2])
	column = as.integer(as.character(cleanup_list[i,3]))
	resp_attr[,column] = gsubkeep(pattern=pattern,replacement=replacement,x=resp_attr[,column])
}

# 2e:
# Assign mean columns for the 4 response categories
resp_answers$DP_Avg = round(apply(X=resp_answers[,2:6],MARGIN=1,FUN=mean),5)
resp_answers$AIP_Avg = round(apply(X=resp_answers[,7:21],MARGIN=1,FUN=mean),5)
resp_answers$GP_Avg = round(apply(X=resp_answers[,22:41],MARGIN=1,FUN=mean),5)
resp_answers$SWLS_Avg = round(apply(X=resp_answers[,42:46],MARGIN=1,FUN=mean),5)

```

## Question 3: HDI table scrape from Wikipedia

```{r Question3 HDI table scrape from Wikipedia}

# Scrape the Wikipedia website for the 8 HDI data tables

URL = 'https://en.wikipedia.org/wiki/List_of_countries_by_Human_Development_Index#Complete_list_of_countries'

html_resp = read_html(URL)

#HDI_data is a single data frame country and HDI
HDI_data = 
	html_nodes(html_resp,css='table.wikitable') %>%
	html_table(fill=TRUE) %>%
	list_rowbind(elements=c(1:8),cols=c(3,4),colname=c('Country','HDI'))

#Cleaning up HDI data and adding categories
HDI_data = subset(HDI_data,HDI_data[,1] != 'Change in rank from previous year[1]')
rownames(HDI_data) = 1:dim(HDI_data)[1]
HDI_data$HDI_Cat = c(rep('Very High',51),rep('High',56),rep('Medium',41),rep('Low',41))
HDI_data$HDI = as.numeric(HDI_data$HDI)

#Attach HDI data to attribute table, and maintain original row ordering
resp_attrwHDI = merge(x=resp_attr,y=HDI_data,by.x='Country_Res',by.y='Country',all.x=TRUE,incomparables=NA)
resp_attrwHDI = resp_attrwHDI[order(resp_attrwHDI$ID),]

```

# Question 4: Preliminary Analysis

```{r Question4 Preliminary Analysis}

# 4A:
# Remove respondents under 18 from resp_attr and resp_answers
row_keep = which(resp_attrwHDI$Age>=18)
resp_attrwHDI = resp_attrwHDI[row_keep,]
resp_answers = resp_answers[row_keep,]

#Merge attributes with question answers for analysis
resp_full = merge(x=resp_answers,y=resp_attrwHDI,by.x='ID',by.y='ID',all.y=TRUE)

# 4B:
#statistics table
stargazer(resp_full[,c('Age','Annual_Inc','HDI','DP_Avg','AIP_Avg','GP_Avg','SWLS_Avg')],type='text',
					summary.stat=c('n','min','p25','median','p75','max','mean','sd'))

#Create 4 plots for sample figure:

h1 = gghistogram(resp_full,x='Age',y='..count..',
								 bins=10,
								 fill='blue',
								 size=2,
								 title='Age',
								 color='white'
)

h1 + font('x.text',size=8)

h2 = gghistogram(resp_full,x='Annual_Inc',y='..count..',
								 bins=10,
								 fill='blueviolet',
								 size=2,
								 title='Ann Inc',
								 color='white'
)

h2 + font('x.text',size=10)

h3 = gghistogram(resp_full,x='HDI',y='..count..',
								 bins=5,
								 fill='cadetblue3',
								 size=2,
								 title='HDI',
								 color='white'
)

DPMetric <- resp_full["DP_Avg"]
DPMetric$Variable <- 'DP_Avg'
colnames(DPMetric) <- c("Value","Variable")

AIPMetric <- resp_full["AIP_Avg"]
AIPMetric$Variable <- 'AIP_Avg'
colnames(AIPMetric) <- c("Value","Variable")

GPMetric <- resp_full["GP_Avg"]
GPMetric$Variable <- 'GP_Avg'
colnames(GPMetric) <- c("Value","Variable")

SWLSMetric <- resp_full["SWLS_Avg"]
SWLSMetric$Variable <- 'SWLS_Avg'
colnames(SWLSMetric) <- c("Value","Variable")

metrics <- rbind(DPMetric, AIPMetric, GPMetric, SWLSMetric)

h4 = ggboxplot(metrics,x='Variable',
							 y='Value',
							 color='Variable',
							 palette='jco',
							 size=1.5
)

h4 + font('x.text',size=10)

ggarrange(ggarrange(h1,h2,h3,ncol=3,labels=c('1','2','3')),
					h4,
					nrow=2,
					labels='4')

# Neither the Age nor Annual Income histograms are particularly normally distributed.  The Age histogram is a bit of a bimodal
# distribution with peaks around the ages of 25, 45 and 55. THe Annual Income is more right-skewed.

# 4C:

#Categorize occupations for table

jobs = cbind(as.character(resp_full$Current_Occ), NA)

jobs_matrix = unlist(apply(jobs,MARGIN=2,FUN=function(x){strsplit(x,' ')})) %>%
	gsub(pattern='[[:punct:]]',replacement='') %>%
	gsub(pattern='^$|^ $',replacement=NA)

jobs_matrix = unlist(apply(jobs,MARGIN=2,FUN=function(x){strsplit(x,' ')})) %>%
	gsub(pattern='[[:punct:]]',replacement='') %>%
	gsub(pattern='^$|^ $',replacement=NA) %>%
	subset(!is.na(jobs_matrix))

jobs_matrix = as.factor(tolower(jobs_matrix))

jobs_count = as.data.frame(table(jobs_matrix))
colnames(jobs_count)=c('title','count')
jobs_count = jobs_count[order(jobs_count$count,jobs_count$title,decreasing=TRUE),]

#Get top 10 job keywords
keywords = jobs_count[1:10,]

keywords_del = c('and','of')
top10Jobs = subset(keywords,!keywords$title %in% keywords_del)

# Frequencies per column 
# Unique distributions
# Gender
apply(resp_full["Gender"], 2, table)

# WorkStatus
apply(resp_full["Work_Status"], 2, table)

#Distributions across Gender and Work Status
with(resp_full, table(Gender, Work_Status))

top10Jobs

# 4D:

# Counts of participants per country in descending order

countryCt <- count(resp_full, resp_full$Country_Res)
colnames(countryCt) <- c("Country", "Count")

sortedCC <- countryCt[order(countryCt$Count, decreasing = TRUE),]

print(sortedCC)

# 4E:

# Find the number of people whose self view matched others view

resp_full$Others_Score <- as.character(resp_full$Others_Score)
resp_full$Self_Score <- as.character(resp_full$Self_Score)

viewMatch <- count(resp_full[which(resp_full$Self_Score == resp_full$Others_Score),])
print(viewMatch)

```

## Question 5: Visualization and Deeper Analysis

```{r Question5 Preliminary Analysis}

# 5B:

DPMeanSub <- resp_full[,c("Country_Res","DP_Avg", "HDI_Cat")]
DPMeanAvg <- aggregate(DP_Avg~Country_Res, DPMeanSub, mean)
DPMeanUnique <- aggregate(HDI_Cat~Country_Res, DPMeanSub, unique)

DPMeanAvgHDI <- merge(DPMeanAvg, DPMeanUnique , by.x=c("Country_Res"),
											by.y=c("Country_Res"))

SortedDP <- DPMeanAvgHDI[order(DPMeanAvgHDI$DP_Avg, decreasing = TRUE),]
Top15DP <- SortedDP[1:15,]

##Set chunk size for output
ggplot(Top15DP, aes(reorder(Country_Res, -DP_Avg), DP_Avg))+
	geom_bar(aes(fill=HDI_Cat), stat="identity")+
	ggtitle("Countries with the top 15 DPMean") +
	theme(plot.title = element_text(hjust = 0.5)) +
	xlab("Country Name")+
	ylab("DP Mean Value")

####START HERE:

# 5C:
AIPMeanSub <- resp_full[,c("Country_Res","AIP_Avg", "HDI_Cat")]
AIPMeanAvg <- aggregate(AIP_Avg~Country_Res, AIPMeanSub, mean)
AIPMeanUnique <- aggregate(HDI_Cat~Country_Res, AIPMeanSub, unique)

AIPMeanAvgHDI <- merge(AIPMeanAvg, AIPMeanUnique , by.x=c("Country_Res"),
											 by.y=c("Country_Res"))

SortedAIP <- AIPMeanAvgHDI[order(AIPMeanAvgHDI$AIP_Avg, decreasing = TRUE),]
Top15AIP <- SortedAIP[1:15,]

##Set chunk size for output
ggplot(Top15AIP, aes(reorder(Country_Res, -AIP_Avg), AIP_Avg))+
	geom_bar(aes(fill=HDI_Cat), stat="identity")+
	ggtitle("Countries with the top 15 AIPMean") +
	theme(plot.title = element_text(hjust = 0.5)) +
	xlab("Country Name")+
	ylab("AIP Mean Value")

#Find countries that exist in both the top 15 AIP and DP Means
merge(Top15DP, Top15AIP, by.x=c("Country_Res"), by.y=c("Country_Res"))

# There are 6 countries that exist in both DP and API top 15 Means.  They are:
# Dominican Republic, Ecuador, Panama, Qatar, Sri Lanka, Uraguay

# 5D:

# Subset merged data frame to only include those rows where Gender is not empty and Annual Income is not empty
mDFMF <- subset(resp_full[c("Country_Res","Age","Annual_Inc","Gender")], Gender != "" & Annual_Inc != "")

# Get the average income per age
mDFMFAvg <- aggregate(Annual_Inc~Age + Gender, mDFMF, mean)

par(mar=c(4,6,2,2))
plot(mDFMFAvg$Annual_Inc ~ mDFMFAvg$Age, xaxt="n", yaxt="n", pch = 16, cex = 1.3, col = ifelse(mDFMFAvg$Gender == "Female", "hot pink", "blue"), xlab="", ylab="")
axis(1,at=pretty(mDFMFAvg$Age),labels=pretty(mDFMFAvg$Age),las=1)
axis(2,at=pretty(mDFMFAvg$Annual_Inc),labels=format(pretty(mDFMFAvg$Annual_Inc),big.mark=",", scientific=FALSE),las=1)
mtext(text="Age", side=1, line=2)
mtext(text="Annual Income", side=2, line=5)
title("Mean Income by Age and Gender")
abline(lm(mDFMFAvg$Annual_Inc ~ mDFMFAvg$Age), col = "green")

# There does seem to be a relationship between age and annual income, with annual income increasing with age.  However,
# there also seems to be a large disparity between males and females and their annual income starting at around age 40 and 
# growing from there.

# 5E:

# Subset merged data frame to only include those rows where Gender is not empty
mDFHDI <- subset(resp_full[c("Country_Res","SWLS_Avg", "HDI", "HDI_Cat")])

mDFHDIAvg <- aggregate(SWLS_Avg~HDI + HDI_Cat, mDFHDI, mean)

par(mar=c(4,6,2,2))
plot(mDFHDIAvg$SWLS_Avg ~ mDFHDIAvg$HDI, xaxt="n", yaxt="n", pch = 16, cex = 1.3, col = "blue", xlab="", ylab="")
axis(1,at=pretty(mDFHDIAvg$HDI),labels=pretty(mDFHDIAvg$HDI),las=1)
axis(2,at=pretty(mDFHDIAvg$SWLS_Avg),labels=format(pretty(mDFHDIAvg$SWLS_Avg),big.mark=",", scientific=FALSE),las=1)
mtext(text="HDI", side=1, line=2)
mtext(text="Mean Life Satisfaction", side=2, line=5)
title("Life Satisfaction Mean by HDI")
abline(lm(mDFHDIAvg$SWLS_Avg ~ mDFHDIAvg$HDI), col = "green")

# Barchart between HDI Category and Mean Life Satisfaction

ggplot(mDFHDIAvg, aes(fill=HDI_Cat, y=SWLS_Avg, x=HDI_Cat)) + 
	scale_x_discrete(labels=c("High Human Development" = "HHD", "Very High Human Development" = "VHHD",
														"Medium Human Development" = "MHD", "Low Human Development" = "LHD")) +
	geom_bar(position="dodge", stat="identity") +
	ggtitle("Mean Life Satisfaction per HDI Category Test") +
	theme(plot.title = element_text(hjust = 0.5)) +
	xlab("Human Development Category") +
	ylab("Life Satisfaction Mean")

```